---
title: "ex3"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

## (1)

\begin{align*}
E[\hat{\beta}_2|\boldsymbol{X}] &= E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{r}|\boldsymbol{X}] \\
&= E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'(\boldsymbol{I}-\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1')\boldsymbol{y}|\boldsymbol{X}] \\
&= E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{y}-(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{y}|\boldsymbol{X}] \\
&= E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{y}|\boldsymbol{X}]-E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{y}|\boldsymbol{X}] \\
&= E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'(\boldsymbol{X}_2\beta_2 + \boldsymbol{X}_1\beta_1 + \varepsilon)|\boldsymbol{X}]-E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{y}|\boldsymbol{X}] \\
&= \beta_2 + (\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1\beta_1-E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{y}|\boldsymbol{X}] \\
&= \beta_2 + (\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1\beta_1-E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'(\boldsymbol{X}_1\beta_1 + \boldsymbol{X}_2\beta_2 + \varepsilon)|\boldsymbol{X}] \\
&= \beta_2 + (\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1\beta_1-E[(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'(\boldsymbol{X}_1\beta_1 + \boldsymbol{X}_2\beta_2 + \varepsilon)|\boldsymbol{X}] \\
&= \beta_2 + (\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1\beta_1-(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1\beta_1 - (\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{X}_2\beta_2 \\
&= \beta_2  - (\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{X}_2\beta_2.
\end{align*}

$$bias~of~\hat{\beta}_2=-(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{X}_2\beta_2.$$ 

## (2)

The OLS solution is $\boldsymbol{b}_2 = (\boldsymbol{X}_2'\boldsymbol{R}_1\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{y}$, where $\boldsymbol{R}_1 = \boldsymbol{I} - \boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'$.

# Question 2

## (1)

$$\boldsymbol{y} = \boldsymbol{X}_1\beta_1+\boldsymbol{X}_2\beta_2+\varepsilon=\boldsymbol{X}_2\boldsymbol{B}\beta_1+\boldsymbol{X}_2\beta_2+\varepsilon=\boldsymbol{X}_2(\boldsymbol{B}\beta_1+\beta_2) + \varepsilon.$$

Hence, $\gamma = \boldsymbol{B}\beta_1 + \beta_2$ can be estimated by regessing $\boldsymbol{y}$ on $\boldsymbol{X}_2$.

## (2)

The RSS of regressing $\boldsymbol{y}$ on $\boldsymbol{X}_2$ is

\begin{align*}
\boldsymbol{e}_2'\boldsymbol{e}_2&=(\boldsymbol{y}-\boldsymbol{X}_2(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{y})'(\boldsymbol{y}-\boldsymbol{X}_2(\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{y})\\
&=\boldsymbol{y}'\boldsymbol{y} - \boldsymbol{b}_2'\boldsymbol{X}_2'\boldsymbol{X}_2\boldsymbol{b}_2,
\end{align*}

where $\boldsymbol{b}_2 = (\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{y}$.


The RSS of regressing $\boldsymbol{y}$ on $\boldsymbol{X}_1$ is

\begin{align*}
\boldsymbol{e}_1'\boldsymbol{e}_1&=(\boldsymbol{y}-\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{y})'(\boldsymbol{y}-\boldsymbol{X}_1(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{y})\\
&=\boldsymbol{y}'\boldsymbol{y} - \boldsymbol{b}_1'\boldsymbol{X}_1'\boldsymbol{X}_1\boldsymbol{b}_1,
\end{align*}

where $\boldsymbol{b}_1 = (\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{y}$.

If $\boldsymbol{B}^{-1}$ exists, then

$$\boldsymbol{b}_2 = (\boldsymbol{X}_2'\boldsymbol{X}_2)^{-1}\boldsymbol{X}_2'\boldsymbol{y} = \boldsymbol{B}(\boldsymbol{B}'\boldsymbol{X}_2'\boldsymbol{X}_2\boldsymbol{B})^{-1}\boldsymbol{B}'\boldsymbol{X}_2'\boldsymbol{y} = \boldsymbol{B}(\boldsymbol{X}_1'\boldsymbol{X}_1)^{-1}\boldsymbol{X}_1'\boldsymbol{y} = \boldsymbol{B}\boldsymbol{b}_1.$$

Hence,

$$\boldsymbol{b}_2'\boldsymbol{X}_2'\boldsymbol{X}_2\boldsymbol{b}_2 = \boldsymbol{b}_1'\boldsymbol{B}'\boldsymbol{X}_2'\boldsymbol{X}_2\boldsymbol{B}\boldsymbol{b}_1 = \boldsymbol{b}_1'\boldsymbol{X}_1'\boldsymbol{X}_1\boldsymbol{b}_1.$$

And the $R^2$ will be identical.

# Question 3

## (1)

\begin{align*}
\boldsymbol{A}\boldsymbol{L}\boldsymbol{A}\boldsymbol{L} &= (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}'(\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}')^{-1}\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}'(\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}')^{-1}\boldsymbol{L}\\
&= (\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}'(\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}')^{-1}\boldsymbol{L}\\
&= \boldsymbol{A}\boldsymbol{L}
\end{align*}


\begin{align*}
\boldsymbol{A}\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}'\boldsymbol{A}' &= \boldsymbol{A}\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}'(\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}')^{-T}\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\\
&= \boldsymbol{A}\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}
\end{align*}

If $\boldsymbol{A}$ is symmetric, then

$$(\boldsymbol{L}\boldsymbol{A})' = \boldsymbol{A}'\boldsymbol{L}'=\boldsymbol{A}\boldsymbol{L}'=\boldsymbol{I}.$$
However, 

$\boldsymbol{A}\boldsymbol{L}'=(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}'(\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}')^{-1}\boldsymbol{L}'$ and

$(\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}')^{-1}\boldsymbol{L}'$ is not well-defined.

Hence, $\boldsymbol{A}$ is not symmetric.

$$\boldsymbol{L}'\boldsymbol{A}'=\boldsymbol{L}'(\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}')^{-1}\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}$$

$\boldsymbol{AL}$ is not symmetric.

## (2)

Since $E[\boldsymbol{X}'\varepsilon|\boldsymbol{X}] = 0$,

$$E[\boldsymbol{b}^*|\boldsymbol{X},\boldsymbol{L},\boldsymbol{r}] = E[\boldsymbol{b}+\boldsymbol{A}(\boldsymbol{r}-\boldsymbol{L}\boldsymbol{b})|\boldsymbol{X},\boldsymbol{L},\boldsymbol{r}]=\boldsymbol{\beta}+\boldsymbol{A}(\boldsymbol{r}-\boldsymbol{L}\boldsymbol{\beta}).$$

Suppose $\boldsymbol{r}-\boldsymbol{L}\boldsymbol{\beta} \neq \boldsymbol{0}$. Then, $bias~of~\boldsymbol{b}^* = \boldsymbol{A}(\boldsymbol{r}-\boldsymbol{L}\boldsymbol{\beta}) \neq \boldsymbol{0}$.

## (3)

$$var[\boldsymbol{b}^*|\boldsymbol{X}] = var[\boldsymbol{b}+\boldsymbol{Ar}-\boldsymbol{ALb}|\boldsymbol{X}]=var[\boldsymbol{b}-\boldsymbol{ALb}|\boldsymbol{X}]=var[(\boldsymbol{I}-\boldsymbol{AL})\boldsymbol{b}|\boldsymbol{X}]=(\boldsymbol{I}-\boldsymbol{AL})var[\boldsymbol{b}|\boldsymbol{X}](\boldsymbol{I}-\boldsymbol{AL})'$$

$$(\boldsymbol{I}-\boldsymbol{AL})var[\boldsymbol{b}|\boldsymbol{X}](\boldsymbol{I}-\boldsymbol{AL})' = (\boldsymbol{I}-\boldsymbol{AL})\sigma^2_\varepsilon(\boldsymbol{X}'\boldsymbol{X})^{-1}(\boldsymbol{I}-\boldsymbol{AL})'=\sigma^2_\varepsilon(\boldsymbol{X}'\boldsymbol{X})^{-1}-\sigma^2_\varepsilon\boldsymbol{A}\boldsymbol{L}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{L}'\boldsymbol{A}'.$$

# Question 4

## (1)

### set 1

```{r}
L = matrix(c(0,1,1,1), nrow = 1)
r = matrix(1)

L
```

```{r}
r
```

### set 2

```{r}
L = matrix(c(0,-1,3,0,
             1,0,0,-1),
           nrow = 2, byrow = TRUE)
r = matrix(c(0,0), ncol = 1)

L
```

```{r}
r
```

### set 3

```{r}
L = matrix(c(0,1,-1,0,
             0,0,0,1,
             1,-3,0,0),
           nrow = 3, byrow = TRUE)
r = matrix(c(2,6,0), ncol = 1)

L
```

```{r}
r
```

## (2)

### set 1

$$Y_i=\beta_1  + \beta_2 X_{i,2}+\beta_3 X_{i,3}+\beta_4 X_{i,4}$$

$$Y_i=\beta_1  + \beta_2 X_{i,2}+\beta_3 X_{i,3}+(1-\beta_2-\beta_3) X_{i,4}$$

$$Y_i-X_{i,4}=\beta_1  + \beta_2 (X_{i,2}-X_{i,4})+\beta_3 (X_{i,3}-X_{i,4})$$

### set 2

$$Y_i=\beta_1  + \beta_2 X_{i,2}+\beta_3 X_{i,3}+\beta_4 X_{i,4}$$

$$Y_i=\beta_1  + 3\beta_3 X_{i,2}+\beta_3 X_{i,3}+\beta_1 X_{i,4}$$

$$Y_i=\beta_1 (1+X_{i,4})  + \beta_3 3X_{i,2}+\beta_3 X_{i,3}$$

### set 3

$$Y_i=\beta_1  + \beta_2 X_{i,2}+\beta_3 X_{i,3}+\beta_4 X_{i,4}$$

$$Y_i=3\beta_2  + \beta_2 X_{i,2}+(\beta_2-2) X_{i,3}+6 X_{i,4}$$
$$Y_i+2 X_{i,3}-6 X_{i,4}=\beta_2 (3+X_{i,2}+X_{i,3})$$




# Question 5














